### 2015-12-28 src/2_Finance.tex
Avance capitulo 2
ordenando propuesta 2
previo rev LSalinas

High frequency trading (HFT) strategies are based on buy and sell assets in
short periods of time gaining a small profit in every transaction, the key is
the amount of transactions that algorithms are capable to execute obtaining profit considering transaction costs. 
In 2012 more than 50\% of US market share of exchange trades were HFT.  HFT has
motivated computer-driven strategies capable of processing large amount of data
in short periods of time.

\section{High Frequency Trading}

HFT is not a strategy but a technology which allow to implement particular
trading strategies. The aim of HFT is to benefit from market short-term pricing
inefficiencies~\cite{chlistalla2011}. HFT is characterised for the use of
high-speed and sophisticated quantitative and algorithmic computer applications
for modelling and executing orders efficiently. In order to make fast decisions,
HFT firms require speed access to trading center servers and sometimes they are
physically near minimising network latencies.

High frequency trades are short-term positions that commonly end the trading
day avoiding leaving opened positions to the next business day. HFT is
frequently associate to algorithm trading strategies, but the former is focused
in to reduce the market impact of large institutional positions and the later refers to trade execution strategies that are typically used by fund managers to buy or sell large amounts of assets. The duration of the positions in HFT it is not well-defined. Figure~\ref{fig:HFTtimes} shows a survey done to traders about the holding time
of a position opened. Most of them agreed that HFT refers to positions between
1-second and 10-minutes. Overnight positions are avoided since they are more
risky and also have more expensive fees, HFT firms end the trading day closing all
opened positions.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=\textwidth]{img/HFTtradestime}
  \caption{Holding time of an opened position of a high frequency trade.}
  \label{fig:HFTtimes}
\end{figure}
In 2014 HFT represented nearly 50\% of equity trades in the US and more than 20\%
in Europe showing a consistent fall since 2009 which was its best year. US market Revenues
have fallen dramatically.
Figure~\ref{fig:HFTmarket} on the left shows HFT trades percentage of US and Europe equity shares. The right figure shows how revenues in US stocks have fallen the last years.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=\textwidth]{img/HFTmarket}
  \caption{Left figure shows HFT market share in the US and Europe. Right figure
  shows revenue in the US.}
  \label{fig:HFTmarket}
\end{figure}
One of the reasons of this fall is that exchange markets have adapted, having
now faster and more transparent and efficient market structures than before.
This has been possible due to its investment in technology enhancing
reliability and stability of transactions. Even though this fall, HFT is still a major component of regulated markets and
will probably remain a topic of interest for researchers in the near future.

@@

A financial market is any marketplace where buyers and sellers participate
trading different assets such as equities, bonds, currencies and derivatives
(future or options). One of the main objectives of financial markets is to set
prices for global trade.
A financial market has many components but the most commonly used are money
markets and capital markets. Money markets are used for short-term basis,
usually for assets up to one year, for greater periods capital markets are used.
Capital markets include the stock or equity market and the bond or debt market
and it their movements are the most widely followed.

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/capitalmarketsnow}
  \caption{Actual structure of capital markets.}
  \label{fig:capitalmarketnow}
\end{figure}
Equity market and foreign exchange market (Forex) are the most popular markets
for high frequency trading strategies. In equity markets can be traded stocks
such as futures and options, exchange-traded funds (ETFs) among others financial
instruments.

Forex market allows the trading of interest rates denominated pair currencies
such as EURUSD (Euro to UD dollar) or USDJPY (US dollar to Japanese yen). Traders of this market are diverse, some of them are
high frequency traders, long-term investors and corporations. The Forex market
used to be centralised, only commercial banks had exclusive access to
inter-dealer networks. Today, Forex market is decentralised and has become the
biggest market in terms of volume of trading. The main participants are
international banks geographically dispersed with continuous 24 hours operations
excepting weekends. Figure~\ref{fig:Forextimes} shows different market trading
hours in GMT for London, New York, Sydney and Tokyo.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/forex-trading-hours}
  \caption{Forex market trading hours (GMT).}
  \label{fig:Forextimes}
\end{figure}

@@

\begin{description}
\item[Market order:] allows to buy or sell an asset at the current best price
whatever the price is. The transaction price is determined by the market
considering previous executed orders and volume requested.
\item[Limit order:] allow to specify the price you want to buy or sell an asset.
Depending on the market conditions, this type of orders can never be executed.
\item[Stop order:] are suitable for investors who are unable to monitor their
investments for a period of time. Stop order allows to specify the price that a
position should be closed called stop price. When the stop price is reached a
market order is executed, this means that market price won't be exactly the same
specified in the stop order.
\end{description}

@@

\begin{description}
\item[Fill or Kill:] is an order that must be executed immediately or being
cancelled, no partial fulfillments are allowed. 
\item[Day:] the order is only valid during the day.
\item[Good til canceled:] a order is active until the investor decides to cancel
it or the trade is executed.
\end{description}

@@

The strong form of the EMH additionally claims that prices instantly reflect even hidden, private or "insider" information.

@@

If the stock market efficiently digests all available information, there is little justification for seeking excess returns gains from investing. However, EMH doesn't lessen the the importance of investing only change its philosophy. The only way an investor can possibly obtain higher returns is by purchasing riskier investments.

>>>>

\vspace{0.5cm}

High frequency trading (HFT) strategies are based on buy and sell assets in
short periods of time gaining a small profit in every transaction despite the transaction costs. The key is
the amount of transactions that HFT algorithms are capable to execute. 
In 2012 more than 50\% of US market trades were HFT.  HFT has
motivated computer-driven strategies capable of processing large amount of data
in short periods of time.

\section{High Frequency Trading Definition}

HFT is not a strategy but a technology that implements different
trading strategies. The aim of HFT is to benefit from market short-term pricing
inefficiencies~\cite{chlistalla2011}. HFT is characterised for the use of
high-speed and sophisticated quantitative and algorithmic computer applications
for modelling and executing orders efficiently. In order to make fast decisions,
HFT firms require speed access to trading servers and sometimes they are
physically near so they can minimise network latencies.

High frequency trades are short-term positions that commonly end the trading
day avoiding leaving opened positions to the next business day. HFT is
frequently associated to algorithm trading strategies, but the former is focused
in to reduce the market impact of large institutional positions and the later refers to trade execution strategies that are typically used by fund managers to buy or sell large amounts of assets. The duration of the positions in HFT it is not well-defined. Figure~\ref{fig:HFTtimes} shows a survey done to traders about the holding time
of a position opened. Most of them agreed that HFT refers to positions between
1-second and 10-minutes. Overnight positions are avoided since they are
riskier and also have more expensive fees, HFT firms end the trading day closing all
opened positions.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=\textwidth]{img/HFTtradestime}
  \caption{Holding time of an opened position of a high frequency trade.}
  \label{fig:HFTtimes}
\end{figure}
In 2014 HFT represented nearly 50\% of equity trades in the US and more than 20\%
in Europe showing a consistent fall since 2009 which was its best year. US market revenues
have fallen dramatically.
Figure~\ref{fig:HFTmarket} on the left shows HFT trades percentage of US and Europe equity shares. The right figure shows how revenues in US stocks have fallen the last years.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=\textwidth]{img/HFTmarket}
  \caption{Left figure shows HFT market share in the US and Europe. Right figure
  shows revenue in the US.}
  \label{fig:HFTmarket}
\end{figure}
One of the reasons of this fall is that exchange markets have adapted, having
now faster, more transparent and efficient market structures than before.
This has been possible due to its investment in technology enhancing
reliability and stability of transactions. Even though this fall, HFT is still a major component of regulated markets and
will probably remain as a topic of interest for researchers in the near future.

@@

A financial market is any marketplace where buyers and sellers participate
trading different assets such as equities, bonds, currencies and derivatives
(future or options). One of the main objectives of financial markets is to set
prices for global trade.
A financial market has many components but the most commonly used are money
markets and capital markets. Money markets are used for short-term basis,
usually for assets up to one year, for greater periods capital markets are used.
Capital markets include the stock or equity market and the bond or debt market
and their movements are the most widely followed.

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/capitalmarketsnow}
  \caption{Actual structure of capital markets.}
  \label{fig:capitalmarketnow}
\end{figure}

Equity market and foreign exchange market (Forex) are the most popular markets
for high frequency trading strategies. In the Equity market, stocks
such as futures and options, exchange-traded funds (ETFs) among others financial
instruments can be traded. Additionally, in the Forex market, interest rates denominated pair currencies 
such as EURUSD (Euro to UD dollar) or USDJPY (US dollar to Japanese yen) are traded. Traders of this market are diverse, some of them are
high frequency traders, long-term investors and corporations. The Forex market
used to be centralised, only commercial banks had exclusive access to
inter-dealer networks. Today, Forex market is decentralised and has become the
biggest market in terms of volume of trading. The main participants are
international banks geographically dispersed with continuous 24 hours operations
excepting weekends. Figure~\ref{fig:Forextimes} shows different market trading
hours in GMT for London, New York, Sydney and Tokyo.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/forex-trading-hours.png}
  \caption{Forex market trading hours (GMT).}
  \label{fig:Forextimes}
\end{figure}

There are three periods of overlaps between different trading times:

\begin{itemize}
\item New York and London from 13:00 GMT until 17:00 GMT
\item Tokyo and London from 8:00 GMT until 9:00 GMT
\item Sydney and Tokyo from 23:00 GMT to 7:00 GMT 
\end{itemize}

These overlap hours are very important since the highest volume of trades are expected. Therefore, most HFT strategies are executed in these hours.

@@

\begin{description}
\item[Market order] allows to buy or sell an asset at the current best price
whatever the price is. The transaction price is determined by the market
considering previous executed orders and volume requested.
\item[Limit order] allow to specify the price you want to buy or sell an asset.
Depending on the market conditions, this type of orders could never been executed.
\item[Stop order] are suitable for investors who are unable to monitor their
investments for a period of time. Stop order allows to specify the price that a
position should be closed called stop price. When the stop price is reached a
market order is executed, this means that market price could not be exactly the same
specified in the stop order.
\end{description}

@@

\begin{description}
\item[Fill or Kill] is an order that must be executed immediately or being
cancelled, no partial fulfillments are allowed. 
\item[Day] the order is only valid during the day.
\item[Good til canceled] a order is active until the investor decides to cancel
it or the trade is executed.
\end{description}

@@

The strong form of the EMH additionally claims that prices instantly reflect even hidden, private or ``insider" information.

@@

If the stock market efficiently digests all available information, there is little justification for seeking excess returns gains from investing. However, EMH doesn't lessen the importance of investing only change its philosophy. The only way an investor can possibly obtain higher returns is by purchasing riskier investments.
